# Selecting and Filtering Data

::: rmdtip
Let's practice importing data, from \@ref(dataframe). Download the [periodic table](https://downgit.github.io/#/home?url=https://github.com/kaisamng/RGuides/blob/main/resources/data/periodic-table-data.csv){target="_blank"} data now, and import it into R as `periodic_table`.

-   Remember to move this into your working directory: `getwd()`

-   Since this is a CSV file, use the **From Text (readr)** function.

-   Change the name of the data set to **periodic_table** before you import it. Remember that's on the bottom left.

:::

The first 10 rows of `periodic_table` are reproduced below.

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(readr)
periodic_table <- read_csv("resources/data/periodic-table-data.csv")
```

```{r, echo=FALSE, fig.cap="The Periodic Table"}
knitr::kable(periodic_table[1:10,], caption="The Periodic Table", booktabs=TRUE) %>%
  scroll_box(height="500px") %>%
  kable_styling(position= "center", font_size = 10)
```

So we have a big data frame of information about the periodic table, but what if we wanted to extract smaller pieces of the data frame?


## Selecting a column using `$`

By default, R will organize character strings in alphabetical order. To see this, we'll introduce two new features: the `table` function and the `$` operator.

Let's say I'm interested in elements in a gas, liquid, and solid state at room temperature. Looking at the `periodic_table` columns, I see that the column that gives me that information is `state_at_stp`.

```{r, attr.output='style="max-height: 100px; font-size:10px;"'}
periodic_table$state_at_stp

```

As you can see, to access specific variables inside a data frame, we can do so by entering the name of the data frame followed by `$` and the name of the variable. (Note that spaces in variable names will not work. You'll likely learn that the hard way, as I have.)

## Selecting an element using `[ , ]` 

Recall in \@ref(vector-selection) section that we could use the `[ ]` operator, to select elements from a single vector. We can extend these operations into a dataframe.

A single vector can be thought of as a dataframe with one row and $n$ columns, which is why you just needed a single number to select an element. Now, since there are multiple rows and columns, you need a comma to differentiate the row and column index.

The `[ ]` operator allows you to select elements within a vector or a dataframe **based on its row and column position.** It starts by locating the row position, and then the column position.

### Selecting a single element using `[ , ]`

Let's say I wanted to select the `name` for `Sodium`, that is in row 11, column 3.

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
periodic_table[11, 3]

periodic_table$name[11]
```

Notice that both commands above give me the desired value-- Sodium-- but the formats are different. + In the first command, R gives me back a table with 1 row and 1 column. This isn't a coincidence, because you pulled directly from the table itself. It's the difference between copying a "cell" in Excel versus double-clicking on the cell, and then copying the text inside. Here, you directly pulled from the table, so R returns back a value within a table. + The second gives me a single character vector. This makes sense-- the `$` operator selects a single vector `name` from the table, and then selects the 11th element. Generally, this is the way you want to select a single table.

Recall the use of `[ ]` when a vector was printed, to help us better understand where we were in printing a large vector. We can use this same tool to select the tenth to the twentieth elements of the `periodic_table$name` variable:

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
periodic_table$name[10:20]
```

## Selecting rows and columns using `[ , ]`

You have now seen how to select specific elements of a vector or a variable, but what if we want a subset of the values in the full data frame across both rows (observations) and columns (variables).?

We can use `[ , ]` where the spot before the comma corresponds to rows and the spot after the comma corresponds to columns. Let's select rows 40 to 50 and columns 1, 2, and 4 from `periodic_table`:

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
periodic_table[41:50, c(1, 2, 4)]
```

### "Let's eat Grandma!" vs. "Let's eat, Grandma!"

::: rmdwarning
**The comma is very important.** Let's look at some examples:
:::

```{r,attr.output='style="max-height: 150px; font-size:10px;"'}
#Returns all of column 2:
periodic_table[2]

#Returns all of row 2:
periodic_table[2, ]

#Returns all of column 2:
periodic_table[ ,2]

#Returns a single element in Row 2, Column 2:
periodic_table[2,2]
```



## Replacing and Removing Data
```{r, echo=FALSE}
friends <- data.frame(names = c("Abram", "Bryant", "Colleen", "David", "Esther", "Jeremiah"),
           ages = c(34, 35, 32, 29, 30, 30),
           DC_Resident = c(TRUE, FALSE, FALSE, TRUE, FALSE, TRUE),
           fav_number = c(1, 2.17, 26, 7, 10, 9),
           stringsAsFactors = FALSE)
```
Destructive operations are any operation where you override the original data. This is not necessairly a bad thing-- sometimes, we don't have a need for the original data, but want to manipulate it in some way. However, it means that if you made a mistake, the only way to revert the mistake is to completely reload the data from its original source, and repeat all your operations again. As in \@ref(vector-operations), you can alter your dataframe by using the storage operator `<-` to overwrite data.

### Replacing one element 

Let's say that I accidentally typed 'Abram' as my friend's name, and I want to replace it with 'Abe'

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
friends$names[1]<- "Abe"
friends
```

### Replacing columns with a mathematical formula

I want to see what everybody's age is in 50 years.

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
friends$ages<- friends$ages+50
friends
```

### Replacing entire columns with another vector

I want to add their last names.

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
friends$names <- c("Abram A", "Bryant B", "Colleen C", "David D", "Esther E", "Jeremiah J")
friends
```

Note in the last example, I had to type the entire vector I planned to override. That's because these are characters, and if you use the `+` operator, R will return an error saying that you can't add "a" with "b". We will discuss a better way to do this, below.

### Dropping Rows

Or I added the first person by mistake to my Friends list, so I have to drop the first row that represents "Abram A":

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
friends<- friends[-1, ]

friends
```



```{=html}
<script>
  iFrameResize({}, ".interactive");
</script>
```

