# Filtering Data

::: rmdtip
Let's practice importing data, from \@ref(dataframe). Download the [periodic table](https://downgit.github.io/#/home?url=https://github.com/kaisamng/RGuides/blob/main/resources/data/periodic-table-data.csv) data now, and import it into R as `periodic_table`.

-   Remember to move this into your working directory: `getwd()`

-   Since this is a CSV file, use the **From Text (readr)** function.

-   Change the name of the data set to **periodic_table** before you import it. Remember that's on the bottom left.

`periodic_table` is reproduced below.
:::

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(readr)
periodic_table <- read_csv("resources/data/periodic-table-data.csv")
```

```{r, echo=FALSE, fig.cap="The Periodic Table"}
knitr::kable(periodic_table, caption="The Periodic Table", booktabs=TRUE) %>%
  scroll_box(height="500px") %>%
  kable_styling(position= "center", font_size = 10)
```

So we have a big data frame of information about the periodic table, but what if we wanted to extract smaller pieces of the data frame?

## Selecting a column using `$`

By default, R will organize character strings in alphabetical order. To see this, we'll introduce two new features: the `table` function and the `$` operator.

Let's say I'm interested in elements in a gas, liquid, and solid state at room temperature. Looking at the `periodic_table` columns, I see that the column that gives me that information is `state_at_stp`.

```{r, attr.output='style="max-height: 100px;"'}
periodic_table$state_at_stp

```

As you can see, to access specific variables inside a data frame, we can do so by entering the name of the data frame followed by `$` and the name of the variable. (Note that spaces in variable names will not work. You'll likely learn that the hard way, as I have.)


## Using `[ , ]` in a dataframe

Recall in \@ref(vector-selection) section that we could use the `[ ]` operator, to select elements from a single vector. We can extend these operations into a dataframe.

A single vector can be thought of as a dataframe with one row and $n$ columns, which is why you just needed a single number to select an element. Now, since there are multiple rows and columns, you need a comma to differentiate the row and column index.

The `[ ]` operator allows you to select elements within a vector or a dataframe **based on its row and column position.** It starts by locating the row position, and then the column position.

### Selecting a single element using `[ , ]`

Let's say I wanted to select the `name` for `Sodium`, that is in row 11, column 3.

```{r}
periodic_table[11, 3]

periodic_table$name[11]
```

Notice that both commands above give me the desired value-- Sodium-- but the formats are different. + In the first command, R gives me back a table with 1 row and 1 column. This isn't a coincidence, because you pulled directly from the table itself. It's the difference between copying a "cell" in Excel versus double-clicking on the cell, and then copying the text inside. Here, you directly pulled from the table, so R returns back a value within a table. + The second gives me a single character vector. This makes sense-- the `$` operator selects a single vector `name` from the table, and then selects the 11th element. Generally, this is the way you want to select a single table.

Recall the use of `[ ]` when a vector was printed, to help us better understand where we were in printing a large vector. We can use this same tool to select the tenth to the twentieth elements of the `periodic_table$name` variable:

```{r}
periodic_table$name[10:20]
```

### Selecting rows and columns using `[ , ]` 

You have now seen how to select specific elements of a vector or a variable, but what if we want a subset of the values in the full data frame across both rows (observations) and columns (variables).?

We can use `[ , ]` where the spot before the comma corresponds to rows and the spot after the comma corresponds to columns. Let's select rows 40 to 50 and columns 1, 2, and 4 from `periodic_table`:

```{r}
periodic_table[41:50, c(1, 2, 4)]
```

### "Let's eat Grandma!" vs. "Let's eat, Grandma!"

::: rmdwarning
**The comma is very important.** Let's look at some examples:
:::

```{r,attr.output='style="max-height: 140px;"'}
#Returns all of column 2:
periodic_table[2]

#Returns all of row 2:
periodic_table[2, ]

#Returns all of column 2:
periodic_table[ ,2]

#Returns a single element in Row 2, Column 2:
periodic_table[2,2]
```

### Filtering data using Logical Vectors

Recall in \@ref(vector-logical-selection) that we could use a logical vector, and the `%in%` operator, to select elements from a single vector. We can extend these operations into a dataframe.

At its most basic level, you can use the `==` operator to test whether one element **exactly** matches another. For example, `"gas"=="gas"` returns TRUE, but `"gas"=="Gas"` will return false. 

Let's say that I want to find any elements that are naturally gasses. 

```{r, attr.output='style="max-height: 200px;"'}
periodic_table[periodic_table$state_at_stp == "Liquid", ]
```

You'll notice that this returns back a massive table of `NAs`, rather than just the two elements that exist in liquid state! Why? 

Let's examine the command `periodic_table$state_at_stp=="Liquid"` more carefully.

```{r, attr.output='style="max-height: 100px;"'}
periodic_table$state_at_stp=="Liquid"
```

  + Every time the element's state is not `Liquid`, R returns False, instead of ignoring the row, which means you get a logical vector of all 118 elements. 
  + This gets passed onto the subset function, which, rather than ignore any row that says FALSE, returns each row as entirely NAs.
  
Although this technically does the job you want, it is not a very elegant solution, which is why the `%in%` operator is better.

```{r}
periodic_table[periodic_table$state_at_stp %in% "Liquid", ]
```
Much better. Notice that, with this call, we didn't specify any columns for R to filter out, so it left those blank. 

Now we can think about how to subset an entire data frame using the same sort of creation of two logical vectors (one for rows and one for columns):

```{r}
periodic_table[ (periodic_table$name %in% c("Hydrogen", "Oxygen") ),
        c("atomic_weight", "state_at_stp")]
```

The extra parentheses around `periodic_table$name %in% c("Hydrogen", "Oxygen")` are a good habit to get into as they ensure everything before the comma is used to select specific rows matching that condition. For the columns, we can specify a vector of the column names to focus on only those variables. The resulting table here gives the `atomic_weight` and `state_at_stp` for `"Hydrogen"` and then for `"Oxygen"`.

There are many more complicated ways to subset a data frame and one can use the `subset` function built into R, but in my experience, whenever you want to do anything more complicated than what we have done here, it is easier to use the `filter` and `select` functions in the `dplyr` package.