# Replacing or Removing Data
```{r, echo=FALSE}
friends <- data.frame(names = c("Abram", "Bryant", "Colleen", "David", "Esther", "Jeremiah"),
           ages = c(34, 35, 32, 29, 30, 30),
           DC_Resident = c(TRUE, FALSE, FALSE, TRUE, FALSE, TRUE),
           fav_number = c(1, 2.17, 26, 7, 10, 9),
           stringsAsFactors = FALSE)
```
Destructive operations are any operation where you override the original data. This is not necessairly a bad thing-- sometimes, we don't have a need for the original data, but want to manipulate it in some way. However, it means that if you made a mistake, the only way to revert the mistake is to completely reload the data from its original source, and repeat all your operations again. As in \@ref(vector-operations), you can alter your dataframe by using the storage operator `<-` to overwrite data.

## Replacing one element 

Let's say that I accidentally typed 'Abram' as my friend's name, and I want to replace it with 'Abe'

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
friends$names[1]<- "Abe"
friends
```

## Replacing columns with a mathematical formula

I want to see what everybody's age is in 50 years.

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
friends$ages<- friends$ages+50
friends
```

## Replacing entire columns with another vector

I want to add their last names.

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
friends$names <- c("Abram A", "Bryant B", "Colleen C", "David D", "Esther E", "Jeremiah J")
friends
```

Note in the last example, I had to type the entire vector I planned to override. That's because these are characters, and if you use the `+` operator, R will return an error saying that you can't add "a" with "b". We will discuss a better way to do this, below.

## Dropping Rows

Or I added the first person by mistake to my Friends list, so I have to drop the first row that represents "Abram A":

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
friends<- friends[-1, ]

friends
```





