# Summarizing Data

Let's do some more practice. Download the [`dca_weather` dataset](https://github.com/kaisamng/RGuides/blob/main/resources/data/dca_weather.csv), which gives the minimum and maximum temperatures in Arlington, VA, and also the amount of rainfall each day.  This data is taken from [NOAA's Website](https://www.ncdc.noaa.gov/cdo-web/).

```{r, echo=FALSE, fig.cap="Weather Data for DCA Airport from 1/1/2022 to 6/19/2022"}
load("~/Github/RGuides/resources/data/dca_weather.RData")
knitr::kable(dca_weather, caption="Weather Data for DCA Airport from 1/1/2022 to 6/19/2022", booktabs=TRUE) %>%
  scroll_box(height="400px") %>%
  kable_styling(position= "center", font_size = 10)
```

## `mean`, `median`, and `fivenum`

These functions do exactly as they sound, however, you must pass a **numeric vector** through each of them. 

```{r, error=TRUE}
#This will return an error-- since dca_weather has multiple columns, R doesn't know which column you want.
mean(dca_weather) 
```

```{r}
mean(dca_weather$TAVG)
```

Knowing the average temperature from January to June is pretty useless though-- sometimes, you want to filter it by month.

```{r}
mean(dca_weather$TAVG[dca_weather$month %in% 6] )
```
If the command looks intimidating, consider the command from inside out.

  1. In the square brackets, you consider only the rows where `dca_weather$months` is equal to 6. Those pass on a set of rows to `dca_weather$TAVG`. 
  2. Then, you pass only those rows, and select the `dca_weather$TAVG` data from those rows. That passes on a vector to `mean()`.
  3. Finally, you calculate the mean of all those numbers. 

The `median()` and `fivenum()` commands work the same. 

```{r}
fivenum(dca_weather$TMIN[dca_weather$month %in% 2])

median(dca_weather$TMAX[dca_weather$month %in% 5])
```


## `sd`




## The `table` function

Let's say I'm interested in how many elements are in gas, liquid, and solid state at room temperature. Looking at the `periodic_table` columns, I see that the column that gives me that information is `state_at_stp`.

```{r}
table(periodic_table$state_at_stp)
```

-   The table function provides a count of the number of elements that appear in each state. But as you can see, the ordering is off.