# Selecting Data Using Logical Conditions

## Using `%in%`

Recall in \@ref(vector-logical-selection) that we could use a logical vector, and the `%in%` operator, to select elements from a single vector. We can extend these operations into a dataframe.

At its most basic level, you can use the `==` operator to test whether one element **exactly** matches another. For example, `"gas"=="gas"` returns TRUE, but `"gas"=="Gas"` will return false.

Let's say that I want to find any elements that are naturally gasses.

```{r, attr.output='style="max-height: 200px; font-size:10px;"'}
periodic_table[periodic_table$state_at_stp == "Liquid", ]
```

You'll notice that this returns back a massive table of `NAs`, rather than just the two elements that exist in liquid state! Why?

Let's examine the command `periodic_table$state_at_stp=="Liquid"` more carefully.

```{r, attr.output='style="max-height: 100px; font-size:10px;"'}
periodic_table$state_at_stp=="Liquid"
```

-   Every time the element's state is not `Liquid`, R returns False, instead of ignoring the row, which means you get a logical vector of all 118 elements.
-   This gets passed onto the subset function, which, rather than ignore any row that says FALSE, returns each row as entirely NAs.

Although this technically does the job you want, it is not a very elegant solution, which is why the `%in%` operator is better.

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
periodic_table[periodic_table$state_at_stp %in% "Liquid", ]
```

Much better. Notice that, with this call, we didn't specify any columns for R to filter out, so it left those blank.

Now we can think about how to subset an entire data frame using the same sort of creation of two logical vectors (one for rows and one for columns):

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
periodic_table[ (periodic_table$name %in% c("Hydrogen", "Oxygen") ),
        c("atomic_weight", "state_at_stp")]
```

The extra parentheses around `periodic_table$name %in% c("Hydrogen", "Oxygen")` are a good habit to get into as they ensure everything before the comma is used to select specific rows matching that condition. For the columns, we can specify a vector of the column names to focus on only those variables. The resulting table here gives the `atomic_weight` and `state_at_stp` for `"Hydrogen"` and then for `"Oxygen"`.

There are many more complicated ways to subset a data frame and one can use the `subset` function built into R, but in my experience, whenever you want to do anything more complicated than what we have done here, it is easier to use the `filter` and `select` functions in the `dplyr` package.

## Using Comparison Operators `<` and `>`
These operators are self explanatory. The `<` and `>` can be used to select rows based on a **numeric** condition.

`!=` is the **not equals** operator. This operator allows for both numeric and character types. 

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
# Prints any row whose atomic number is less than 10.
periodic_table[periodic_table$atomic_number<10,]

#Prints any row whose state at standard room temperature is not Solid.
periodic_table[periodic_table$state_at_stp != "Solid",]
```

## Combining logical operators with "or", "and" conditions
Finally, the `&` operator means **and**, and the `|` operator means **or**. You can fund `|` by pressing the shirt key and kitting the "reverse slash" key above the `enter` key on your keyboard. 

It allows you to combine multiple logical statements together, which may be useful if you have multiple criteria to work on.

```{r, attr.output='style="max-height: 150px; font-size:10px;"'}
#Prints any element which is Non-metal AND whose  atomic number is less than 15
periodic_table[periodic_table$description != "Non-metal" & periodic_table$atomic_number <15,]

#Prints any element which is Non-metal OR whose atomic number is less than 15
periodic_table[periodic_table$description != "Non-metal" | periodic_table$atomic_number <15,]
```

# Vectors and Dataframe Exercises {-}

<iframe style="margin:0 auto; min-width: 100%;" id="ch2-4-exercises" class="interactive" src="https://kaisamng.shinyapps.io/Chapters_2_4_Exercises/" scrolling="no" frameborder="no">

</iframe>


<script>
  iFrameResize({}, ".interactive");
</script>

