---
title: "Two-Sample T-Test"
author: "Mr. Ng"
output: 
  word_document: 
    reference_docx: style.docx
    toc: yes
    fig_width: 8
    fig_caption: yes
    df_print: tibble

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fontsize= 10, mainfont= "Calibri")

```

## Start by Loading the Dataset from a CSV File. 

The data that I've downloaded is currently in a CSV file. I'm going to use the function File>> Import Dataset >> from base (readr) option. 

When it first gets loaded, I want certain columns to be treated as numeric vectors. The code is below. 

```{r read animecsv}
library(readr)
anime <- read_csv("anime.csv", col_types = cols(MAL_ID = col_number(), Score = col_number(), Episodes = col_number(), Popularity = col_number(), Members = col_number(), Favorites = col_number(), Watching = col_number(), Completed = col_number(), `On-Hold` = col_number(), Dropped = col_number(),`Plan to Watch` = col_number(), `Score-10` = col_number(), `Score-9` = col_number(), `Score-8` = col_number(), `Score-7` = col_number(), `Score-6` = col_number(), `Score-5` = col_number(), `Score-4` = col_number(), `Score-3` = col_number(), `Score-2` = col_number(),`Score-1` = col_number()))
```

## Data Wrangling

Examine any problems with the dataset. Notice that there are a lot of randome problems where we treated a column as a numeric vector, but the actual data value was "Unknown" instead. R will convert the value to "NA," which is fine with us.

```{r problems}
problems(anime)
View(anime)
names(anime)
```

If you look, for some reason the "Genres" column is named "Genders." Let's rename to "Genres." 
```{r rename column}
names(anime)[names(anime)=="Genders"] <- "Genres"
```
Verify that the column names are now correct.
```{r rename, echo=FALSE}
names(anime)
```

Save the new file as a R dataframe.
```{r save RData}
save(anime, file="anime.RData")
```

## Isolate the variables of interest.

This uses subsets all rows within the anime dataframe based on the condition in the brackets, and then saves it into a new dataframe called action_anime. The condition is satifised with the "grep" function, which will search for any character string and return "TRUE" if it contains "Action" within the column anime$Genres. 

```{r subset action anime}
action_anime<- anime[grep("Action", anime$Genres), ]
```

View the dataframe. How many animes do you recognize? 
```{r view action anime}
View(action_anime)
head(action_anime)
```

We're interested in how people rated these animes. Let's calculate the score column for each. 

```{r calculate mean}
mean(action_anime$Score, na.rm=TRUE)
mean(anime$Score, na.rm=TRUE)
```


Interesting! Let's isolate the score columns from both dataframes and store them as separate vectors. 
```{r isolate score vectors}
anime_score<- anime$Score
action_anime_score <- action_anime$Score
```


## Graph boxplots.

It looks like the average rating for action animes is indeed higher than the average rating for all animes. But by how much? To answer this question, let's first graph a boxplot.
```{r boxplot}
boxplot(anime_score, action_anime_score, 
        ylim=c(0,10),
        main= "Boxplot comparison of All Anime vs. Action Anime",
        xlab= "Rating",
        names=c("Anime", "Action Anime"),
        col=c("Red", "Green"),
        horizontal=TRUE)
```

## Run a Two Sample T-Test.

Based on the boxplots, it looks like there's a difference. Action animes appear to be rated higher than the rest. That means:

* Our null hypothesis is that there is no difference between the two averages, so u1 - u2 =0.

* Our alternative hypothesis is that there **is** a difference between two averages, specifically that the average rating for all animes is lower than the average rating for action animes. 

Set the alternative hypothesis to be that the true average score of all animes is LOWER than the true average score for action animes. 

```{r t_test}
action_t_test <- t.test(anime_score, action_anime_score, alternative= "less", var.equal=TRUE)
```
View the results. 

```{r view t test, echo=FALSE}
action_t_test
```

Wow! There's a serious difference. Look at the p-value for that. Now we need to graph this. 

## Graph the t-distribution and test statistic. 

Let's start by:

1. storing the degrees of freedom in a variable called action_t_test.

2. creating an x-axis and store it as a variable called **t_d_st_x_axis**. This is the equivalent of drawing tick marks everywhere on a graph, but a ton of tick marks. We're making a huge sequence 10^4 numbers long (so 10000 numbers) between -15.5 and 15.5. This helps fit in my massive test statistic of -14.34

```{r}
action_t_test_df<- length(anime_score) +length(action_anime_score) -2 #Recall how to calculate degrees of freedom.
t_dist_x_axis<- seq(-15.5, 15.5, length= 10^4)
```

Finally, plot. Notice here that for plot():

* the x-value is **t_dist_x_axis**

* the y-value is **dt(t_dist_x_axis, df=action_t_test_df)**. This generates a t-value for every value in **t_dist_x_axis**, so basically *f(x)*, with the degrees of greedom we set earlier. 

```{r t-test plot, eval=FALSE}
plot(t_dist_x_axis, dt(t_dist_x_axis, df=action_t_test_df), 
     type='l', 
     xlab='t', 
     ylab='',
     main="The most ridiculous t-test graph ever",
     sub="Look at that test statistic in the dotted line. Just look.")
```

I want to show how ridiculous this is. We're going to store the critical t* value of 0.01 in it's own variable, and plot that with red lines. 
```{r critcal t lines, eval=FALSE}
crit_t_value_0.01<- qt(0.01, df=action_t_test_df)
abline(v=crit_t_value_0.01, col='red')
abline(v=-crit_t_value_0.01, col='red')
```



Finally, plot the test statistic, using a dotted line. 

```{r line for test statistic, eval=FALSE}
abline(v=action_t_test$statistic, lty=2)
```

```{r t-test plot2, echo=FALSE}
plot(t_dist_x_axis, dt(t_dist_x_axis, df=action_t_test_df), 
     type='l', 
     xlab='t', 
     ylab='',
     main="The most ridiculous t-test graph ever",
     sub="Look at that test statistic in the dotted line. Just look.")
crit_t_value_0.01<- qt(0.01, df=action_t_test_df)
abline(v=crit_t_value_0.01, col='red')
abline(v=-crit_t_value_0.01, col='red')
abline(v=action_t_test$statistic, lty=2)
```









